---
title: "R Notebook"
output: html_notebook
---

```{r}
library(car)
library(MASS)
library(plotrix)
```


# Exercise 2
## a)
```{r}
ker <- read.table("http://www-eio.upc.es/~jan/Data/MVA/kernels.dat", header=T)
head(ker)
```

## b) Perimeter and area are the most correlated.
```{r}
scatterplotMatrix(ker)
cor(ker)
```

## c) Seems that 2 components are enough.
```{r}
ker2 <- scale(ker)
acp <- princomp(ker2)
summary(acp)
screeplot(acp)
```

## d) The second component is principally afected by the compactness. While the first one is almost a mean, except for compactness. It can be seen as the second component almost being compactness and the first the rest.
```{r}
biplot(acp, cex=0.5)
```

## e) Lenght and groove, it isn't the same as the one given by the correlation matrix

## f) It is greater in all components
```{r}
apply(acp$scores, 2, sd)
acp$sdev
```

## g) It gives the projection matrix.
```{r}
acp$loadings
```

## h)
```{r}
plot(acp$scores[,1], acp$scores[,2], xlab="First component", ylab="Second component", pch=17, col="red")
text(acp$scores[,1]+0.2, acp$scores[,2]+0.2, labels=rownames(ker))
```


## i) The correlation between different principal components is almost 0, as expected, i.e. the variables are independent.

```{r}
cor(as.matrix(acp$scores))
```


## j)

```{r}
X <- scale(ker)
Y = svd(X)
print(Y$d)
```

## k)

```{r}
pc <- Y$u%*%diag(Y$d)
head(pc)
```


## l)

```{r}
plot(pc[,1:2], xlab="", ylab="")
arrows(x0=rep(0,7),y0=rep(0,7),x1=t(Y$v[1,1:7]),y1=t(Y$v[2,1:7]), length = 0.1, col="red")
draw.circle(0,0,radius=1)
```

## m)

```{r}
(abs(pc[1,1])+abs(pc[1,2]))/sum(abs(pc[1,]))
```

## n)

```{r}
sum(Y$d)
```

## o)

```{r}
sum(Y$d[1:3])/sum(Y$d) * 100
```

## p) Basically represents the asymmetry variable.

```{r}
Y$v
```

# 4
## a)

```{r}
dist <- as.matrix(read.table("http://www-eio.upc.es/~jan/data/MVA/GeneticDist.txt"))
dist[1:6,1:6]

raza <- as.factor(c(rep("Afro", 342), rep("Caucasico", 361), rep("Asiatico", 97)))
head(raza)
```

## b)

```{r}
pco <- cmdscale(dist, k = 3, eig=T)
plot(pco$points[,1], pco$points[,2], xlab="", ylab="", col=c(rep("tan4", 342), rep("bisque", 361), rep("yellow", 97)))
legend(x=-0.2, y=-0.1, legend=c("Afro", "Caucasian", "Asiatic"), fill=c("tan4","bisque","yellow"), cex=0.5)
```

## c) Yes, if k is equal to the number of genetics variables. Or k = 800 too.

## d) Yes, because the distance matrix is centered to compute the MDS.

```{r}
sum(abs(pco$eig)<1E-13)
```

## e)

```{r}
sum(abs(pco$eig[1:2]))/sum(abs(pco$eig))
```

## f) It doesn't help, in fact it mixes two races.

```{r}
plot(pco$points[,1], pco$points[,3], xlab="", ylab="", col=c(rep("tan4", 342), rep("bisque", 361), rep("yellow", 97)))
legend(x=-0.2, y=-0.1, legend=c("Afro", "Caucasian", "Asiatic"), fill=c("tan4","bisque","yellow"), cex=0.5)

sum(abs(pco$eig[c(1,3)]))/sum(abs(pco$eig))
```

## g) It has a lot of variance and 

```{r}
d2 <- as.matrix(dist(pco$points[,1:2]))
plot(x=dist[lower.tri(dist)], y=d2[lower.tri(d2)], xlab="observed", ylab="predicted", type="p")
abline(a=0,b=1, col=red)
```

## h) 

```{r}
set.seed(1234)
n <- 800
init <- scale(matrix(runif(n*2), ncol=2), scale=c(F,F))

nMDS <- isoMDS(dist, k = 2, y = init)

plot(nMDS$points[,1], nMDS$points[,2], xlab="", ylab="", col=c(rep("tan4", 342), rep("bisque", 361), rep("yellow", 97)))
legend(x=-0.2, y=-0.1, legend=c("Afro", "Caucasian", "Asiatic"), fill=c("tan4","bisque","yellow"), cex=0.5)

nMDS$stress
```


